---
title: "Julia’s Tab 2 Attempt"
author: "Julia Dagum"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Join energy generation plant info with locational plant info. Filtered for Solar PV.

```{r}
# ----------
# Install ibraries
# ---------

library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(sf)

```

```{r}

# ----------
# Read in data 
# ----------

gen_by_plant <- read_csv(here::here("data", "annual_generation_by_power_plant.csv"))

plant_info <- read_csv(here::here("data", "plant_info.csv"), 
                       skip = 12, col_names = TRUE)

# --------
# Clean up data
# --------

tidy_gen_by_plant <- gen_by_plant %>% 
  clean_names() %>% 
  select(year, state, cec_plant_id, status, start_date, retire_date, prime_mover_id, prime_mover_desc, capacity, gross_m_wh, net_m_wh) %>% 
  rename("gross_mwh" = "gross_m_wh") %>% 
  rename("net_mwh" = "net_m_wh") %>% 
  # filter for just photovoltaics (this excludes solar thermal, wind, etc.)
  # but there are still plants that say the energy source is natural gas/other non-solar...
  filter(prime_mover_id == "PV")
  # for visuals, remember to filter out non-operational systems once they have a "retire date" so they still show up if a year is selected before retirement

tidy_plant_info <- plant_info %>% 
  clean_names() %>% 
  # we'll already have state in the other df
  select(-state)

# --------
# Join dataframes
# ---------

plants_location_join <- inner_join(tidy_plant_info, 
                                   tidy_gen_by_plant, 
                                   "cec_plant_id")
# note that there are a 19 entries wiht energy_source_category not equal to "SUN" but prime_mover_id equal to "PV"

```
Some entries have primary energy source as natural gas or biogas, which we're assuming is the energy input needed to run the plant. The ones with SUN as primary energy source have slight losses between gross mwh and net mwh. We'll use net mwh to measure capacity. 

Prime Mover ID = The type of engine or turbine that is the source of power for driving the generator
Primary Energy Source = Primary fuel type used by the generating unit. 

### Get spatial data

```{r}

# Read in CA counties
ca_counties <- read_sf(dsn = here::here("data", "ca_counties"), 
                       layer = "california_county_shape_file") %>% 
  rename(county = "NAME") %>% 
  select(county) %>% 
  group_by(county) %>% 
  summarize()

# Get number of plants per county from plants_location_join

plants_per_county <- plants_location_join %>% 
  group_by(county) %>% 
  tally()
# tally counts number of plants in each county

# Join plant information to CA counties and make sf object with sticky geometries
plant_county_join <- full_join(plants_per_county, ca_counties, "county")
plant_county_sf <- st_as_sf(plant_county_join)


```

```{r}
# -------------------------------------
# I don't actually think it makes sense to join cities from cal enviroscreen:
# 
# # Use nearby cities from Cal Enviro-screen data to join with plants
# # Read in data set
# env_scrn_df <- read_csv(here::here("data", "CalEnviroScreen.csv")) %>% 
#   clean_names() %>% 
#   rename(nearby_city = nearby_city_to_help_approximate_location_only,
#          lon = longitude,
#          lat = latitude,
#          county = california_county) %>% 
#   select(-contains("pctl"), -contains("percentile"))
# 
# # Convert to sf
# env_scrn_sf <- st_as_sf(env_scrn_df, coords = c("lon", "lat"), crs = 4326) %>% 
#   rename(city = "nearby_city")
# 
# # Try joining plant cities with cal enviro-screen nearby cities
# plant_city_join <- full_join(plants_location_join, env_scrn_sf, "city")

```

### Make map with solar plants by county
*Not sure if this is the right number because some of the plants have prime_mover_id as photovotaic but then the energy_source_category column is not solar...
```{r}
# Does it make sense to use st_intersection (from session 3 video)??
# Not sure what that would do

# Instead just use plant_county_sf

# Make map of CA with plants per county
# widget will be filter by year
# put reactive df in what waldo commented out
# temp_plot and placeholder
# Could try to do this using solar capacity in MW per county??
ggplot() +
  geom_sf(data = plant_county_sf, 
          aes(fill = n), 
          size = 0.2, 
          color = "white") +
  scale_fill_continuous(low = "yellow", high = "red")
```

